---
title: "SV_PCA_analysis"
author: "Julija Cvetocnika"
date: "05/06/2023"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(DESeq2)
library(sva)
library(tximport)
library(AnnotationDbi)


```
#Sets up tx2gene (allows you to map from transcripts to genes)
```{r}
setwd("/media/julija/Shield2")
txdb <- makeTxDbFromGFF("Mus_musculus.GRCm39.105.gtf")
k <- keys(txdb, keytype ="TXNAME")
tx2gene <- select(txdb, k, "GENEID", "TXNAME")
```
#Reads in biotypes file. This lists all protein-coding genes, which are the only ones we're interested for this analysis.
```{r}
setwd("/media/julija/Shield2")
biotypes <- read.delim("biotypes.txt", sep = " ")
```

#Filters tx2gene based on protein-coding genes only
```{r}
setwd("/media/julija/Shield2")
protein_coding <- biotypes[biotypes$gene_biotype == "protein_coding",]
tx2gene_filt <- tx2gene[tx2gene$GENEID %in% protein_coding$ensembl_gene_id,]
```


#1: this Sets up path to our files. NOTE: you will have to change the samples file names (with all the groups and info)depending on what your folders are called.

#2:Creates an object called txi.salmon, this is essentially all of your gene count files side by side.

```{r}
setwd("/media/julija/Shield2/Align/male2")
samples <- read.delim("SampleData_M_SIvG_Hipp.txt", header = TRUE, sep = "\t")
samples$replicate <- factor(samples$replicate)

files <- file.path(paste0(samples$File_name), "abundance.tsv")
names(files) <- paste0(samples$File_name)
file.exists(files)
head(files)
factor_cols <- sapply(samples, is.factor)

txi.kallisto.tsv.male <- tximport(files, type = "kallisto", tx2gene = tx2gene_filt, ignoreTxVersion = TRUE)
```

#First analysis using DEseq2, this is to get a normalised count matrix which we can use to find surrogate variables using SVAseq. Note the only factor being compared here is Group_Combined.
```{r}

setwd("/media/julija/Shield1/Align/male2")
dds <- DESeqDataSetFromTximport(txi = txi.kallisto.tsv.male,
                                colData = samples, 
                                design = ~ Group_Combined)

keep <- rowSums(counts(dds)) >= 100
dds <- dds[keep,]


dds$Group_Combined <- relevel(dds$Group_Combined, ref = "MVG-Hipp")
dds <- DESeq(dds)
```

#Uses the normalised count matrix to find Number of Surrogate Variables (SVs)
```{r}
counts <- counts_normalized <- counts(dds, normalized=TRUE)

mod <- model.matrix(~ Group_Combined,samples)
mod0 <- model.matrix(~ 1, samples)
svseq = svaseq(counts_normalized,mod,mod0)$sv

```
#Creats a new samples object with the surrogate variables (SVs) attached. NOTE: You will have to vary this based on the number of SVs. i.e. if you only have 2 SVs you'll have "colnames(svseq) <- c("S1", "S2")"

```{r}
colnames(svseq) <- c("S1", "S2")
samples2 <- cbind(samples, svseq)
```

#Final ANalysis with SVs included as covariates. NOTE: you'll have to change the number of covariates if you only have 2 etc.

```{r}
#set up 
setwd("/media/julija/Shield1/Align/male2")
dds <- DESeqDataSetFromTximport(txi = txi.kallisto.tsv.male,
                                colData = samples2, 
                                design = ~ S1 + S2 +  Group_Combined)

keep <- rowSums(counts(dds)) >= 100
dds <- dds[keep,]


dds$Group_Combined <- relevel(dds$Group_Combined, ref = "MVG-Hipp")
dds <- DESeq(dds)
```

#Write Final Analysis into working directory
```{r}
setwd("/media/julija/Shield1/Align/male2")
res <- results(dds, name="Group_Combined_MVSI.Hipp_vs_MVG.Hipp")
summary(res)
write.table(res, "Results.txt")
```



#rlog transform your counts for principal component analysis (PCA)
```{r}
rld <- assay(rlog(dds))
rownames(samples) <- samples$Full.Name
colnames(rld) <- samples$Full.Name
```

#Plot PC1 vs. PC2 and colour by group

```{r}
samples$replicate <- factor(samples$replicate)
pca = PCAtools::pca(rld, metadata = samples)
biplot(pca, colby = "Group_Combined", title = "Male Hipp.")
```

```{r}
function(dds){

	key <- intersect(rownames(dds)[which(dds$log2FoldChange>=1)], rownames(dds)[which(dds$pvalue<=0.05)])

    results_upreg_male <- res((dds)[which(rownames(dds) %in% key),])
	return(results_upreg_male)
}
```

```{r}
framed_fem %>%
  gather(Sample, counts) %>%
  ggplot(aes(Sample, counts)) +
  geom_boxplot()
```

```{r}
plotMA(res, ylim=c(-2,2))
```





