---
title: "Untitled"
output: html_document
date: "2023-06-14"
---

```{r}

data <- read.delim("competitve_ewce_analysis.txt", header = T , sep= "\t")

```



```{r}



```


```{r}
#x what we observe, t what we expect 

#four vectors, one for pvalues, one for enrichment and two for confidence intervals.

pvalues <- c()
enrichment <- c()
confidence1 <- c()
confidence2 <- c()


for (i in 1:24){
  pvalues <- c(pvalues, poisson.test(x=c(data$observedDEGS[i], data$observedBG[i]), T=c(data$expectedDEGS[i], data$expectedBG[i]))$p.value)
  
  enrichment <- c(enrichment, poisson.test(x=c(data$observedDEGS[i], data$observedBG[i]), T=c(data$expectedDEGS[i], data$expectedBG[i]))$estimate)
  
confidence1 <- c(confidence1, poisson.test(x=c(data$observedDEGS[i], data$observedBG[i]), T=c(data$expectedDEGS[i], data$expectedBG[i]))$conf.int[1])

confidence2<- c(confidence2, poisson.test(x=c(data$observedDEGS[i], data$observedBG[i]), T=c(data$expectedDEGS[i], data$expectedBG[i]))$conf.int[2])
                  
}



```

```{r}
alldata <-data 

alldata$p.val <- pvalues
alldata$enrich <- enrichment
alldata$conf_in1 <- confidence1
alldata$conf_in2 <- confidence2
```



number of degs, how many in each 
break it up nto upregulated and down regulated and hwat overlaps between the two - venn diagram 
gene ontology analysis - cpdb- neuronal ontologies sunburt, take top for general and sunburst block for the single 
cell type analysis - table show which cell types in each have corrected p value of 0.05 - top male and femae, box where there is cell type name, what it is and the pvalue
denovolyser -  te two gene sets, number in each and which are significntly enriched, none for me. 

magma - table of p values and beta values for two groups and four diseases. 

syngo - sets of genes that are overrepresented in the genesets. Loads of genes from one ontology, something might have been changed by SI. 
cell type analysis - markers 
de novo - rare mutations in magma - human data - show convergence with human data 
